\chapter{Preface}

This manual was written as a summary and companion guide to the course titled ``Computernetwerken met \abbr{TCP}/\abbr{IP}''\footnote{This title translates to \emph{Computer networks with} \abbr{TCP}/\abbr{IP}.} that I teach at Syntra Bizz.
Although I teach that course in Dutch, this companion is written in English and this for two reasons.
First of all this allows me to reach a wider audience with this manual.
Secondly, it is difficult to write a technology book in Dutch, given that so many terms are in English.
I found it too difficult to decide which technical terms to translate from English to Dutch.
It is impossible to translate every technical term and translating none of them -- which is really the best option so that it would be easier to search for the terms online to find more resources on them -- would make the text a real mix of Dutch and English.
I decided it is better to write the complete text in English.

I teach the accompanying course in three days of six hours each for a total of eightteen hours of in-class training.
Additionally I expect the students to make lab exercises at home.
Should you consider doing these labs in class also, the course would take about a forty-hour work week.
The schedule for the course is given in \vref{tab:course-schedule}.

\begin{table}
   \sffamily
   \centering
   \begin{tabular}{rlll}
                       & \textsc{day 1} & \textsc{day 2}       & \textsc{day 3}   \\[1ex]
   \textit{morning}    & \abbr{IP} addressing  & \abbr{TCP}/\abbr{UDP} and Ethernet & cabling en Wi-Fi  \\
   \textit{afternoon}  & Packet Tracer  & \abbr{VLAN} and \abbr{STP}       & applications     \\
   \end{tabular}
   \caption{In-class course schedule}
   \label{tab:course-schedule}
\end{table}

This manual consists of two parts.
The first part covers the topics that I teach in class.
The second part contains examples and exercises.
I assume the students -- and thus also the reader -- have access to Cisco Packet Tracer but really any virtualisation or simulation software will do.
Other possible candidates include \href{https://www.eve-ng.net/}{eve-ng} and \href{https://www.gns3.com/}{GNS3}.

Most of the examples given are using the Cisco IOS syntax but there are also examples for Juniper Networks' Junos.
Example configuration snippets for other products, such as HP ProCurve or the newer Aruba switches might be added in the future.
%, Aruba's ArubaOS and HP ProCurve.

Last but not least, I welcome constructive feedback at info@marcoen.net.